\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-08 paper/paper.tex
more updates.

These two performance bottlenecks illustrated in Figure~\ref{fig:problem} are
not \emph{independent}. In particular, durability semantics mean that the
key-value store  cannot respond to the client until the operation has been
persisted to the log, since it must be completed even on a crash and subsequent
recovery. Therefore, addressing the network bottleneck is pointless without also
addressing the storage bottleneck. Cyclone shows how a small amount of
non-volatile memory can be used to address both bottlenecks.

@@

\begin{figure}
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Cyclone event handling}
\label{fig:control_plane}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.35]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6]{results2/replicas.pdf}
\caption{Impact of replica count}
\label{fig:replicas}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/512.pdf}
\caption{Impact of payload size}
\label{fig:payload}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog.pdf}
\caption{Impact of adding second level log}
\label{fig:flashlog}
\end{figure}

@@

\begin{figure*}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog_512.pdf}
  \caption{Second level log with 512 bytes payload}
  \label{fig:flashlog_512}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication performance}
  \label{fig:kv_baseline}
\end{minipage}
\end{figure*}

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. A
comparison is hard since Mojim is designed for the hypothetical scenario where
there is enough persistent memory to hold all persistent data, while Cyclone is
designed for the practical case where both the write-ahead log and key value
store data structures need to be maintained largely on flash. Nevertheless we
can compare their replication performance for the case where we use only the NVM
log in Cyclone. On 40Gbps infiniband, Mojim reported being able to mirror data
at 4GB/s. In contrast, Cyclone can replicate data three ways using 4*10 Gbps
commodity ethernet NICs at a peak of roughly 3.7 GB/s. This is with Cyclone is
also running the RAFT consensus protocol, while Mojim is doing simple mirroring
of main memory. One can conclude that network offload via RDMA and Infiniband is
not an a-priori requirement for high throughput replication. This is made
possible by Cyclone's approach that relieves the CPU from the burden of data
copies. On the other hand, the latency of a hop on commodity ethernet at 5 us
with our DPDK stack is larger than the close to 2us latency with
Infiniband. Depending on the use case for the key value store application the
extra 6us with ethernet for the round trip to the replica might be a concern for
some. We note that mapping the commutativity of key value store interfaces to
multiple logs in Cyclone is a technique that can also be applied to RDMA. Mojim
confirms that concurrency is a significant determinant of performance when
replicating over RDMA.

@@

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using the commutativity
of the key value store interface. Although CRANE is more generally applicable in
that it makes no assumptions about the application API, deterministic
multithreading comes with high overheads for applications that store large
amounts of data - CRANE reports around a 2X slowdown for MySQL, likely rendering
it unsuitable as a replacement for the write ahead log in key value stores.

>>>>

These two performance bottlenecks illustrated in Figure~\ref{fig:problem} are
\emph{not independent}, they cannot be addressed in isolation. In particular,
durability semantics mean that the key-value store cannot respond to the client
until the operation has been persisted to the log, since it must be completed
even on a crash and subsequent recovery. Therefore, addressing the network
bottleneck is pointless without also addressing the storage bottleneck. Cyclone
shows how a small amount of non-volatile memory can be used to address both
bottlenecks.

@@

\begin{figure*}
  \begin{minipage}{.2\textwidth}
    \scriptsize
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Event handling}
\label{fig:control_plane}
\end{minipage}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[scale=0.3]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{minipage}
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[scale=0.3]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{minipage}
\end{figure*}

@@

@@

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/replicas.pdf}
  \caption{Impact of replica count}
  \label{fig:replicas}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/512.pdf}
  \caption{Impact of payload size}
  \label{fig:payload}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog.pdf}
  \caption{Adding second level log}
  \label{fig:flashlog}
\end{minipage}
\end{figure*}

@@

@@

\begin{figure*}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog_512.pdf}
  \caption{Second level log - 512 bytes payload}
  \label{fig:flashlog_512}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{minipage}
\end{figure*}

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data
structures. Mojim is a weak replication scheme in that none of the numerous
\emph{evaluated} modes of Mojim provide any means to keep more than two replicas in
strong synchronization. Its multi-tiered replication could lose committed
updates and therefore its guarantees are too weak to be used as a replacement
for RocksDB's write ahead log. Further, Mojim makes the somewhat strange claim
that ``Paxos-like protocols'' require two networking round trips and therefore
are somehow inefficient compared to mirroring. This is not true, as we have
demonstrated with Cyclone where the leader responds as soon as it hears back
from a majority quorum, the critical path to replication therefore being one
network round trip.  A performance comparison with Mojim is hard, since it is
designed for the hypothetical scenario where there is enough persistent memory
to hold all persistent data, while Cyclone is designed for the practical case
where both the write-ahead log and key value store data structures need to be
maintained largely on flash. Nevertheless we can compare their replication
performance for the case where we use only the NVM log in Cyclone. On 40Gbps
infiniband, Mojim reported being able to mirror data at 4GB/s. In contrast,
Cyclone can replicate data three ways using 4*10 Gbps commodity ethernet NICs at
a peak of roughly 3.7 GB/s. This is with Cyclone is also running the RAFT
consensus protocol, while Mojim is doing simple mirroring of main memory. One
can conclude that network offload via RDMA and Infiniband is not an a-priori
requirement for high throughput replication. This is made possible by Cyclone's
approach that relieves the CPU from the burden of data copies. On the other
hand, the latency of a hop on commodity ethernet at 5 us with our DPDK stack is
larger than the close to 2us latency with Infiniband. Depending on the use case
for the key value store application the extra 6us with ethernet for the round
trip to the replica might be a concern for some. We note that mapping the
commutativity of key value store interfaces to multiple logs in Cyclone is a
technique that can also be applied to RDMA. Mojim confirms that concurrency is a
significant determinant of performance when replicating over RDMA.

@@

{CRANE}~\cite{crane} is a system for replicating mutithreaded programs using
Paxos. It uses deterministic multi-threading to ensure replicas converge to the
same state as opposed to semantic equivalence in Cyclone using API
commutativity.  CRANE reports around a 2X slowdown for MySQL, likely rendering
it unsuitable as a replacement for the write ahead log in key value stores.
